---
title: "genreNetwork"
author: "Clint McKenna"
date: "October 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this exercise, we will look at TMDB data (https://www.kaggle.com/tmdb/tmdb-movie-metadata). We will create a network of genres in the dataset, where each genre is a node. The size of the nodes will coorespond to the frequency of genres (larger = more movies). Finally, the edges will indicate that a movie shares two or more genres.


## data prep
```{r}

# load packages
library(igraph)
library(tidyverse)

# functions
select <- dplyr::select

# read in data
movies <- read.csv('../data/clean/movies.csv')
actors <- read.csv('../data/clean/actors.csv')

# just get the first 10 actors listed on each movie
actors <- actors %>%
  #arrange(title) %>%
  group_by(title) %>%
  slice(c(1:5)) %>%
  ungroup()
  

# get actors from highest budget movies
top <- movies %>% 
  select(title, budget) %>%
  #group_by(title) %>%
  top_n(n = 100, wt = budget)
actors <- left_join(top, actors, by = 'title')


# get sample of 500 movies for genre list
# actors <- actors %>%
  # sample(., 500) %>%
  #filter(substr(title, 1, 1) == 'A')
  

# merge back with genres


```


So let's write a function to get the combinations we are looking for.
```{r}

# function for acquiring genre combinations
# x will be the input for the fuction - the dataframe we are using
getCombinations <- function(x) {

  # the column we are looking to make into an edge list is in quotes
  input <- x[,'name']
  
  # here, we make a new object within our function for possible combinations
  combinations <- combn(input, m = 2) %>% 
  t()
  
  # and output a dataframe, which is returned when function is run
  output <- data.frame(
    actor1 = combinations[,1],
    actor2 = combinations[,2]
  )
  
  return(output)
  rm(combinations, output)
}



```





So here is the steps we will take:
- For each movie, calculate the combinations of genres possible
- 


```{r}
# remove films that only have one genre (one row in dataframe)
# here, we group by individual movies, add a column for the genre count
# then, filter out movies where count is 1 or less
actorList <- actors %>%
  group_by(title) %>%
  mutate(count = n()) %>%
  filter(count > 2) %>%
  select(-count) %>% 
  ungroup() %>%
  mutate_all(as.character) %>%
  as.data.frame()

# use split to split up dataframe into individual lists for each movie
actorList <- actorList %>%
  # mutate(group = group_indices(., title)) %>%
  # split(.$group) %>%
  # map_dfr(getCombinations, .id = 'group')
  split(.$title) %>%
  map_dfr(getCombinations, .id = 'title')


```


Now, let's preview edgelist.
```{r}
# preview
head(actorList)
tail(actorList)

```


The last movie, Zulu, should have 3 genres: Crime, Drama, Thriller. Let's check that against the original genres dataset to make sure.
```{r}
actors %>%
  filter(title == 'Avatar')

```

Great! Now with our edgelist, we can proceed to visualizing the network. For this, we will use the igraph package.
```{r}
# for the next steps, we need to move title to the last column
actorList <- actorList %>%
  select(actor1, actor2, title)

# create network object. non-directed
net <- graph_from_data_frame(
  actorList,
  directed = FALSE
)

# plot
plot(net, 
     layout = layout_nicely,
     # vertex.label.cex=0.75,
     # vertex.label.font=2,
     # vertex.label.color='black', 
     vertex.label = NA,
     vertex.shape='circle', 
     vertex.size = 2,
     vertex.color = 'plum',
     
     edge.width = 1
     )

```


Add edge weights
```{r}
E(net)$weight <- 1


simplify(net, edge.attr.comb = list(weight = 'sum'))


plot(net)

```

